---
title: "Introduction to EarthquakeVisualization"
author: "Manuel"
date: "2020-12-17"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{earthquakevisualization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(EarthquakeVisualization)
```

The EarthquakeVisualization R package provides functionality to process, clean
and visualize earthquake datasets downloaded from the [NOAA Significant
Earthquake Database] (https://www.ngdc.noaa.gov/hazel/view/hazards/earthquake/search).

## eq_clean_data
This function applies some cleaning processes to a dataframe originiting from
the \href{https://www.ngdc.noaa.gov/hazel/view/hazards/earthquake/search}{NOAA Significant Earthquake Database}.
Specifically, it converts the Latitude and Longitude columns to numeric class
and combines year, month and day into a single variable in the Date class. The
dataset needs to be read before using this function. Example:
```{r, eval=FALSE}
library(readr)
filename="earthquakes.tsv"
data=read_delim(filename,delim="\t")
data=data[-1,]
cleandata <- eq_clean_data(rawdata)
```

## eq_location_clean
This function applies some cleaning processes to the Location Name column in
a dataset downloaded from the \href{https://www.ngdc.noaa.gov/hazel/view/hazards/earthquake/search}{NOAA Significant Earthquake Database}.
Specifically, it drops any location beyond that or those of the first
occurring country, it drops the preceding name of country and region, and
converts the resulting location name to title case. Example:
```{r, eval=FALSE}
cleandata <- eq_location_clean(data)
```

## geom_timeline
This function implements a ggplot2 geom that allows to display earthquake data
along a time line. Data from multiple countries can be represented in parallel
time lines. Earthquake magnitude, total number of deaths and other variables can
be also shown using typical ggplot2 parameters, such as color, or size. The
dataframe needs to be processed by eq_clean_data before using geom_timeline.
```{r, eval=FALSE}
library(ggplot2)
library(magrittr)
data %>% eq_clean_data() %>% ggplot() +
  geom_timeline(aes(x = Date, y = Country, size = Mag, color = `Total Deaths`)
```


## geom_timeline
This function implements a ggplot2 geom that allows to annotate an earthquake
timeline plot created using geom_timeline. The dataframe needs to be processed
by eq_clean_data and eq_location_clean before using geom_timeline_label.
```{r, eval=FALSE}
library(ggplot2)
library(magrittr)
data %>% eq_clean_data() %>% eq_location_clean() %>% ggplot() +
  geom_timeline(aes(x = Date, y = Country, size = Mag, color = `Total Deaths`) +
  geom_timeline_label(aes(x = Date, y = Country, label = `Location Name`, mag = Mag, n_max = 5))
```


## eq_create_label

This function creates an HTML label as a new column in an earthquake dataframe
where Location, Magnitude and Total Deaths are introduced, together with
labels, and that can be used for Leaflet map annotation. The dataframe needs to
be processed at least by eq_location_clean before using eq_create_label.
```{r, eval=FALSE}
annotateddata <- eq_create_label(data)
```


## eq_map
This function creates an interactive Leaflet map to visualize earthquake
locations contained in the input dataframe marking them with a circle whose
size is proportional to the earthquake's magnitude. Earthquake locations can
be annotated.
```{r, eval=FALSE}
library(magrittr)
data %>% eq_clean_data() %>% eq_location_clean() %>% eq_create_label() +
  eq_map(annot_col="popup_text")
```

